---
title: 'Lab 6: Correlation'
---

```{r, include = FALSE}
# This option turns evaluation of R code off in this document. Remove it
# if you use it as a template for solving the lab exercise.
knitr::opts_chunk$set(eval = FALSE)
```

This session is concerned with summary statistics for interpoint
correlation (i.e. dependence between points).  
The lecturer's R script is [available here](https://raw.githubusercontent.com/spatstat/AAU2015/master/Scripts/script06.R) (right click and save).

```{r include=FALSE}
library(spatstat)
```

### Exercise 1

The `swedishpines` dataset was recorded in a study plot
in a large forest. We shall assume the pattern is stationary.

1.  Calculate the estimate of the $K$-function using `Kest`.

2.  Plot $\widehat K(r)$ against $r$

3.  Plot $\widehat K(r) - \pi r^2$ against $r$.

4.  Calculate the estimate of the $L$-function and plot it against
    $r$.

5.  Plot $\widehat L(r) - r$ against $r$.

6.  Calculate and plot an estimate of the pair correlation function
    using `pcf`.

7.  Draw tentative conclusions from these plots about interpoint
    interaction in the data.


### Exercise 2

1.  Generate Fry Plots for the Swedish Pines data using the two commands
    ```{r}
    fryplot(swedishpines)
    fryplot(swedishpines, width=50)
    ```

2.  What can you interpret from these plots?


### Exercise 3

The `japanesepines` dataset is believed to exhibit
spatial inhomogeneity. The question is whether, after allowing for
inhomogeneity, there is still evidence of interpoint interaction. We
will use the inhomogeneous $K$-function.

1.  Compute the inhomogeneous $K$ function using the default
    estimate of intensity (a leave-one-out kernel smoother) with
    heavy smoothing:
    ```{r}
    KiS <- Kinhom(japanesepines, sigma=0.1)
    plot(KiS)
    ```
	
2.  Fit a parametric trend to the data, and use this to compute the
    inhomogeneous $K$ function:
    ```{r}
    fit <- ppm(japanesepines ~ polynom(x,y,2))
    lambda <- predict(fit, type="trend")
    KiP <- Kinhom(japanesepines, lambda)
    plot(KiP)
    ```
	
3.  Plot corresponding estimates of the inhomogeneous $L$ function,
    using either `Linhom` or
    ```{r}
    plot(KiS, sqrt(./pi) ~ r)
    ```
    and similarly for `KiP`.

4.  Draw tentative conclusions about interpoint interaction.


### Exercise 4

To understand the difficulties with the $K$-function when the point
pattern is not spatially homogeneous, try the following experiment.

1.  Generate a simulated realisation of an inhomogeneous Poisson
    process, e.g.
    ```{r}
    X <- rpoispp(function(x,y){ 200 * exp(-3 * x) })
    ```
	
2.  Plot the $K$-function or $L$-function. It will most likely appear to show
    evidence of clustering.


### Exercise 5

The cell process (`rcell`) has the same theoretical $K$-function as
the uniform Poisson process.

1.  Read the help file

2.  Generate a simulated realisation of the cell process 
    with a 10x10 grid of cells and plot it.
    
3.  Plot the $K$ or $L$-function for this pattern, and
    determine whether it is distinguishable from a Poisson process.
    