---
title: 'Lab 9: Cluster and Cox processes'
---

```{r, include = FALSE}
# This option turns evaluation of R code off in this document. Remove it
# if you use it as a template for solving the lab exercise.
knitr::opts_chunk$set(eval = FALSE)
```

This session is concerned with summary statistics for interpoint
spacing and distances.

```{r include=FALSE}
library(spatstat)
```

### Exercise 1

The command `rThomas` generates simulated realisations of
the Thomas model (‘modified Thomas cluster process’).

1.  Read the help file.

2.  Type `plot(rThomas(10, 0.05, 8))` a few times, and
    interpret the results.

3.  Experiment with the arguments of `rThomas` to obtain
    point patterns that
    1.  consist of a few, well-separated, very tight clusters of
        points;
    2.  look similar to realisations of a uniform Poisson process.


### Exercise 2

1.  Read the help file for `kppm`.

2.  Fit the Thomas model to the `redwood` data by the
    method of minimum contrast:
    ```{r}
    fit <- kppm(redwood ~ 1, clusters="Thomas")
    fit
    plot(fit)
    ```

3.  Read off the parameters of the fitted model, and generate a
    simulated realisation of the fitted model using `rThomas`.

4.  Type `plot(simulate(fit))` to generate a simulated realisation
    of the fitted model automatically.

5.  Try the command
    ```{r}
    fit2 <- kppm(redwood ~ 1, clusters="Thomas", startpar=c(kappa=10, scale=0.1))
    ```
    and briefly explore the fitting algorithm’s sensitivity to the
    initial guesses at the parameter values `kappa` and `scale`.

6.  Generate and plot several simulated realisations of the fitted
    model, to assess whether it is plausible.

7.  Extract and plot the fitted pair correlation function by
    ```{r}
    pcffit <- pcfmodel(fit)
    plot(pcffit, xlim = c(0, 0.3))
    ```

8.  Type `plot(envelope(fit, Lest, nsim=39))` to generate simulation
    envelopes of the $L$ function from this fitted model. Do they
    suggest the model is plausible?


### Exercise 3

1.  Fit a Matern cluster process to the `redwood` data.

2.  Use `vcov` to estimate the covariance matrix of the parameter estimates.

3.  Compare with the covariance matrix obtained when fitting a homogeneous Poisson model.