---
title: 'Lab 8: Spacing and distances'
---

```{r, include = FALSE}
# This option turns evaluation of R code off in this document. Remove it
# if you use it as a template for solving the lab exercise.
knitr::opts_chunk$set(eval = FALSE)
```

This session is concerned with summary statistics for
spacings and interpoint distances.

```{r include=FALSE}
library(spatstat)
```

### Exercise 1

For the `swedishpines` data:

1.  Calculate the estimate of the nearest neighbour distance
    distribution function $G$ using `Gest`.

2.  Plot $\hat G(r)$ against $r$

3.  Plot $\hat G(r)$ against the theoretical (Poisson) value
    $G_{\mbox{pois}}(r) = 1 - \exp(-\lambda \pi r^2)$.

4.  Define Fisher’s variance-stabilising transformation for c.d.f.’s
    by
    ```{r}
    Phi <- function(x) asin(sqrt(x))
    ```
    Plot the $G$ function using the formula `Phi(.) ~ Phi(theo)`
    and interpret it.


### Exercise 2

For the `swedishpines` data:

1.  Calculate the estimate of the nearest neighbour distance
    distribution function $F$ using `Fest`.

2.  Plot $\hat F(r)$ against $r$

3.  Plot $\hat F(r)$ against the theoretical (Poisson) value
    $F_{\mbox{pois}}(r) = 1 - \exp(-\lambda \pi r^2)$.

4.  Define Fisher’s variance-stabilising transformation for c.d.f.’s
    by
    ```{r}
    Phi <- function(x) asin(sqrt(x))
    ```
    Plot the $F$ function using the formula `Phi(.) ~ Phi(theo)`
    and interpret it.
