---
title: 'Lab 3: Intensity dependent on covariate'
---

```{r, include = FALSE}
# This option turns evaluation of R code off in this document. Remove it
# if you use it as a template for solving the lab exercise.
knitr::opts_chunk$set(eval = FALSE)
```

This session covers tools for investigating intensity depending on a covariate.  
The lecturer's R script is [available here](https://raw.githubusercontent.com/spatstat/AAU2015/master/Scripts/script03.R) (right click and save).

```{r include=FALSE}
library(spatstat)
```


### Exercise 1

The `bei` dataset gives the locations of trees in a survey area with additional
covariate information in a list `bei.extra`.

1.  Assign the elevation covariate to a variable `elev` by typing
    ```{r}
    elev <- bei.extra$elev
    ```

2.  Plot the trees on top of an image of the elevation covariate.

3.  Cut the study region into 4 areas according to the value of the 
    terrain elevation, and make a texture plot of the result.

4.	Convert the image from above to a tesselation, count the number 
    of points in each region using `quadratcount`, and plot the
    quadrat counts.

5.  Estimate the intensity in each of the four regions.


### Exercise 2

Assume that the intensity of trees is a function $\lambda(u) = \rho(e(u))$
where $e(u)$ is the terrain elevation at location u.

1.  Compute a nonparametric estimate of the function $\rho$ and plot it by
    ```{r}
    rh <- rhohat(bei, elev)
    plot(rh)
    ```

2.	Compute the predicted intensity based on this estimate of $\rho$.

3.  Compute a non-parametric estimate by kernel smoothing and
  	compare with the predicted intensity above.

4.  Bonus info: To plot the two intensity estimates next to each
  	other you collect the estimates as a spatial object list
  	(`solist`) and plot the result (the estimates are called
  	`pred` and `ker` below):
    ```{r}
    l <- solist(pred, ker)
    plot(l, equal.ribbon = TRUE, main = "", 
         main.panel = c("rhohat prediction", "kernel smoothing"))
    ```


### Exercise 3

Continuing with the dataset `bei` conduct both Berman's Z1 and Z2 tests for
dependence on `elev`, and plot the results.
