---
title: 'Lab 1: Introduction'
---

```{r, include = FALSE}
# This option turns evaluation of R code off in this document. Remove it
# if you use it as a template for solving the lab exercise.
knitr::opts_chunk$set(eval = FALSE)
```

This session is about reading in, displaying and summarising point patterns.

If you have not already done so, you’ll need to start R and load the `spatstat`
package by typing
```r
library(spatstat)
```

### Exercise 1
We will study a dataset that records the locations of Ponderosa Pine
trees (*Pinus ponderosa*) in a study region in the
Klamath National Forest in northern California. The data are
included with `spatstat` as the dataset `ponderosa`.

1.  assign the data to a shorter name, like `X` or `P`;

2.  plot the data;

3.  find out how many trees are recorded;

4.  find the dimensions of the study region;

5.  obtain an estimate of the average intensity of trees (number of
    trees per unit area).

### Exercise 2

The Ponderosa data, continued:

1.  When you type `plot(ponderosa)`, the command that is
    actually executed is `plot.ppp`, the plot method for
    point patterns. Read the help file for the function
    `plot.ppp`, and find out which argument to the
    function can be used to control the main title for the plot;

2.  plot the Ponderosa data with the title *Ponderosa Pine
    Trees* above it;

3.  from your reading of the help file, predict what will happen if
    we type
    ```{r}
    plot(ponderosa, chars="X", cols="green")
    ```
    then check that your guess was correct;

4.  try different values of the argument `chars`, for
    example, one of the integers 0 to 25, or a letter of the
    alphabet. (Note the difference between `chars=3` and
    `chars="+"`, and the difference between `chars=4` and
    `chars="X"`).

### Exercise 3

The following vectors record the locations of 10 scintillation events
observed under a microscope. Coordinates are given in microns, and
the study region was $30 \times 30$ microns, with the origin at the
bottom left corner.
```r
x <- c(13, 15, 27, 17, 8, 8, 1, 14, 19, 23)
y <- c(3, 15, 7, 11, 10, 17, 29, 22, 19, 29)
```
Create a point pattern `X` from the data,
and plot the point pattern (use `owin` or `square` to define the study region).

### Exercise 4

The file `anthills.txt` is available in the Data directory on 
[github](https://github.com/spatstat/AAU2015) and downloadable by 
[this direct link](https://raw.githubusercontent.com/spatstat/AAU2015/labs/Data/anthills.txt) (right click and save).

It records the locations of anthills recorded in a
1200x1500 metre study region in northern Australia.
Coordinates are given in metres, along with a letter code recording
the ecological ‘status’ of each anthill (in this exercise we will ignore this letter code).

1.  read the data into `R` as a data frame, using the `R` function
    `read.table`. (Since the input file has a header line, you
    will need to use the argument `header=TRUE` when you call
    `read.table`.)

2.  check the data for any peculiarities.

3.  create a point pattern `hills` containing these data.
    Ensure that the unit of length is given its correct name.

4.  plot the data.
