---
title: 'Lab 10: Gibbs processes'
---

```{r, include = FALSE}
# This option turns evaluation of R code off in this document. Remove it
# if you use it as a template for solving the lab exercise.
knitr::opts_chunk$set(eval = FALSE)
```

This session is concerned with Gibbs models for point patterns with interpoint interaction.

```{r include=FALSE}
library(spatstat)
```

### Exercise 1

In this question we fit a Strauss point process model to
the `swedishpines` data.

1.  We need a guess at the interaction distance $R$. Compute and
    plot the $L$-function of the dataset and choose the value $r$
    which maximises the discrepancy $|L(r)-r|$.

2.  Fit the stationary Strauss model with the chosen interaction
    distance using
    ```{r}
    ppm(swedishpines, ~1, Strauss(R))
    ```
    where `R` is your chosen value.

    3.  Interpret the printout: how strong is the interaction?

    4.  Plot the fitted pairwise interaction function using
        `plot(fitin(fit))`.

### Exercise 2

In Question 1 we guesstimated the Strauss interaction distance 
parameter. Alternatively this parameter could be estimated by 
profile pseudolikelihood.

1.  Look again at the plot of the $L$-function of
    `swedishpines` and determine a plausible range of
    possible values for the interaction distance.

2.  Generate a sequence of values equally spaced across this range,
    for example, if your range of plausible values was
    $[0.05, 0.3]$, then type
    ```{r}
    rvals <- seq(0.05, 0.3, by=0.01)
    ```

3.  Construct a data frame, with one column named `r`
    (matching the argument name of `Strauss`), containing
    these values. For example
    ```{r}
    D <- data.frame(r = rvals)
    ```

4.  Execute
    ```{r}
    fitp <- profilepl(D, Strauss, swedishpines)
    ```
    to find the maximum profile pseudolikelihood fit.

5.  Print and plot the object `fitp`.

6.  Compare the computed estimate of interaction distance $r$ with
    your guesstimate. Compare the corresponding estimates of the
    Strauss interaction parameter $\gamma$.

7.  Extract the fitted Gibbs point process model from the object
    `fitp` as 
    ```{r}
    bestfit <- as.ppm(fitp)
    ```

### Exercise 3

For the Strauss model fitted in Question 1,

1.  Generate and plot a simulated realisation of the fitted model
    using `simulate`.

2.  Plot the $L$-function of the data along with the global
    simulation envelopes from 19 realisations of the fitted model.


### Exercise 4

1.  Read the help file for `Geyer`.

2.  Fit a stationary Geyer saturation process to `swedishpines`, with the 
    same interaction distance as for the Strauss model computed in Question 2,
    and trying different values of the saturation parameter `sat = 1, 2, 3` say.

3.  Fit the same model with the addition of a log-quadratic trend.

4.  Plot the fitted trend and conditional intensity.

### Exercise 5

Modify question 1 by using the Huang-Ogata approximate
maximum likelihood algorithm (`method="ho"`) instead of maximum
pseudolikelihood (the default, `method="mpl"`).

### Exercise 6

Repeat Question 2 for the inhomogeneous Strauss process with
log-quadratic trend. The corresponding call to `profilepl` is
```{r}
fitp <- profilepl(D, Strauss, swedishpines, ~polynom(x,y,2))
```
    
### Exercise 7

Repeat Question 3 for the inhomogeneous Strauss process with
log-quadratic trend, using the inhomogeneous $L$-function `Linhom`
in place of the usual $L$-function.
